---
output:
  html_document:
    theme:
      bg: '#121212'
      fg: white
  pdf_document: default
---

Import Libraries
```{r}
library(devtools)
library(spotifyr)
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggridges)
library(geniusr)
library(tidytext)
```



Accessing Spotify API
```{r}
Sys.setenv(SPOTIFY_CLIENT_ID = '94671b31e007451a9d8caa8443b46dc8')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '501a6f69715b46679a9bb4a7fa578210')
access_token <- get_spotify_access_token()
```



Read In Playlists Of Top Hits From 2015 through 2022
```{r}
topOf2015 <- get_playlist_audio_features("Spotify", "37i9dQZF1DX9ukdrXQLJGZ")
topOf2015$Year <- 2015
topOf2016 <- get_playlist_audio_features("Spotify", "37i9dQZF1DX8XZ6AUo9R4R")
topOf2016$Year <- 2016
topOf2017 <- get_playlist_audio_features("Spotify", "37i9dQZF1DWTE7dVUebpUW")
topOf2017$Year <- 2017
topOf2018 <- get_playlist_audio_features("Spotify", "37i9dQZF1DXe2bobNYDtW8")
topOf2018$Year <- 2018
topOf2019 <- get_playlist_audio_features("Spotify", "37i9dQZF1DWVRSukIED0e9")
topOf2019$Year <- 2019
topOf2020 <- get_playlist_audio_features("Spotify", "37i9dQZF1DX7Jl5KP2eZaS")
topOf2020$Year <- 2020
topOf2021 <- get_playlist_audio_features("Spotifyn", "5GhQiRkGuqzpWZSE7OU4Se")
topOf2021$Year <- 2021
topOf2022 <- get_playlist_audio_features("Spotify", "37i9dQZF1DXbJMiQ53rTyJ")
topOf2022$Year <- 2022

topOfPastDecade <- rbind(topOf2015, topOf2016, topOf2017, topOf2018, topOf2019, topOf2020, topOf2021, topOf2022)
```



Plot Duration Distribution From 2015 through 2022
```{r}
durationPlotPastDecade <- topOfPastDecade %>%
  ggplot(aes(track.duration_ms/1000, as.factor(Year))) + 
  geom_density_ridges(scale = 3.2,
                      colour=alpha("#121212", 1), rel_min_height = 0.01,
                      size = 2, fill = "#57acdc", alpha = 1) +
  theme(panel.background = element_rect(fill = '#121212', color = '#121212'),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y = element_text(color="#57acdc", 
                                   size=12, family = "Arial Bold"), 
        axis.text.x = element_text(color="#57acdc", 
                                   size=12, family = "Arial Bold"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = , l = 0))) +
  xlab("Duration (seconds)") + ylab("Release Year") + theme(text = element_text(color = "#57acdc", size = 15, family = "Arial Bold"),  plot.margin = margin(1.5, 1.5, 1.5, 1.5, "cm")) +
  theme(plot.background = element_rect(fill = "#121212"), legend.background = element_rect(fill = "#121212") ) 

durationPlotPastDecade
```



Plot Valence Difference Between 2000 and 2022
```{r}
topOf2000 <- get_playlist_audio_features("Spotify", "37i9dQZF1DWUZv12GM5cFk")
topOf2000$Year <- 2000

valence2000_2022 <- rbind(topOf2000, topOf2022)

valencePlotPastDecade <- valence2000_2022 %>%
  ggplot(aes(valence, as.factor(Year))) + 
  geom_density_ridges(scale = 3.2,
                      colour=alpha("#121212", 1), rel_min_height = 0.01,
                      size = 2, fill = "#57Dcbe", alpha = 1) +
  theme(panel.background = element_rect(fill = '#121212', color = '#121212'),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y = element_text(color="#57Dcbe", 
                                   size=12, family = "Arial Bold"), 
        axis.text.x = element_text(color="#57Dcbe", 
                                   size=12, family = "Arial Bold"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = , l = 0))) +
  xlab("Valence") + ylab("Release Year") + theme(text = element_text(color = "#57Dcbe", size = 15, family = "Arial Bold"),  plot.margin = margin(1.5, 1.5, 1.5, 1.5, "cm")) +
  theme(plot.background = element_rect(fill = "#121212"), legend.background = element_rect(fill = "#121212") ) 

valencePlotPastDecade
```




Plot Danceability From 2015 through 2022
```{r}
#valencePlotPastDecade <- valence2000_2022 %>%
danceabilityPlotPastDecade <-topOfPastDecade%>% ggplot(aes(danceability, as.factor(Year))) + 
  geom_density_ridges_gradient(scale = 3.2,
                      colour=alpha("#121212", 1), rel_min_height = 0.01,
                      size = 2, fill = "#E91E63", alpha = 1) +
  theme(panel.background = element_rect(fill = '#121212', color = '#121212'),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y = element_text(color="#E91E63", 
                                   size=12, family = "Arial Bold"), 
        axis.text.x = element_text(color="#E91E63", 
                                   size=12, family = "Arial Bold"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = , l = 0))) +
  xlab("Danceability") + ylab("Release Year") + theme(text = element_text(color = "#E91E63", size = 15, family = "Arial Bold"),  plot.margin = margin(1.5, 1.5, 1.5, 1.5, "cm")) +
  theme(plot.background = element_rect(fill = "#121212"), legend.background = element_rect(fill = "#121212") ) 

danceabilityPlotPastDecade
```




Access Genius API
```{r}
Sys.setenv(GENIUS_API_TOKEN = '_agVKgO1niNZsa_ULiKekTLfQ6tqHOKTfPPJ8VJRYNm0zuUd59BiS063XV5IPYwG')
genius_token()
```




Lyrics Analysis For Billboard Hot 100 2010
```{r}
BillboardHot100 <- get_playlist_audio_features("Billboard", "6UeSakyzhiEt4NB3UAd6NQ")
BillboardHot100$track.name[8] <- "Lift Me Up"


billboardLyrics <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(billboardLyrics) <- c("track.name", "word", "geniusSongID")
curBillboardLyrics <- data.frame()

billboardLyricsFunction <- function(i){
  
  if ( nrow(BillboardHot100[[28]][[i]]) == 1 ) {
    curBillboardlyrics <- get_lyrics_id(as.character((search_song(BillboardHot100$track.name[i]) %>%filter(artist_name == BillboardHot100[[28]][[i]]$name[1]))$song_id[1]))
  }
  
  if ( nrow(BillboardHot100[[28]][[i]]) == 2 ) {
    curBillboardlyrics <-  get_lyrics_id(as.character(((search_song(BillboardHot100$track.name[i]) %>%filter(artist_name == paste(BillboardHot100[[28]][[i]]$name[1], BillboardHot100[[28]][[i]]$name[2], sep = " & " )))$song_id[1]   )))
  }
  
  curBillboardlyrics <- curBillboardlyrics %>% unnest_tokens(word, line)
  data(stop_words)
  curBillboardlyrics <- curBillboardlyrics %>% anti_join(stop_words)
  curBillboardlyrics <- data.frame(track.name = curBillboardlyrics$song_name, word = curBillboardlyrics$word, geniusSongID = curBillboardlyrics$song_id)
  billboardLyrics <- rbind(billboardLyrics, curBillboardlyrics)
  return(billboardLyrics)
}

for (i in 1:100) {
  try(billboardLyrics <- billboardLyricsFunction(i), silent = TRUE)
}

billboardLyrics$track.name <- tolower(billboardLyrics$track.name)
BillboardHot100$track.name <- tolower(BillboardHot100$track.name)

billboardLyrics <- left_join(billboardLyrics, BillboardHot100, by = "track.name")
billboardLyrics <- billboardLyrics %>% inner_join(get_sentiments("bing"))

billboardLyricsSummary <- billboardLyrics %>% summarise(positiveWords = sum(sentiment == "positive"), negativeWords = sum(sentiment == "negative"), posPercentage = positiveWords / (positiveWords + negativeWords), negPercentage = negativeWords / (positiveWords + negativeWords))
head(billboardLyricsSummary)

```




Lyrics Analysis For Billboard Hot 100 2010
```{r}
Top100of2010 <- get_playlist_audio_features("Wicked Dreamer", "4aUY170nZ3mhkzMpTAXDv2")

billboard2010Lyrics <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(billboard2010Lyrics) <- c("track.name", "word", "geniusSongID")
curBillboardLyrics <- data.frame()

billboard2010LyricsFunction <- function(i){
  
  if ( nrow(Top100of2010[[28]][[i]]) == 1 ) {
    curBillboardlyrics <- get_lyrics_id(as.character((search_song(Top100of2010$track.name[i]) %>%filter(artist_name == Top100of2010[[28]][[i]]$name[1]))$song_id[1]))
  }
  
  if ( nrow(Top100of2010[[28]][[i]]) == 2 ) {
    curBillboardlyrics <-  get_lyrics_id(as.character(((search_song(Top100of2010$track.name[i]) %>%filter(artist_name == paste(Top100of2010[[28]][[i]]$name[1], Top100of2010[[28]][[i]]$name[2], sep = " & " )))$song_id[1]   )))
  }
  
  curBillboardlyrics <- curBillboardlyrics %>% unnest_tokens(word, line)
  data(stop_words)
  curBillboardlyrics <- curBillboardlyrics %>% anti_join(stop_words)
  curBillboardlyrics <- data.frame(track.name = curBillboardlyrics$song_name, word = curBillboardlyrics$word, geniusSongID = curBillboardlyrics$song_id)
  billboard2010Lyrics <- rbind(billboard2010Lyrics, curBillboardlyrics)
  return(billboard2010Lyrics)
}

for (i in 1:100) {
  try(billboard2010Lyrics <- billboard2010LyricsFunction(i), silent = TRUE)
}

billboard2010Lyrics$track.name <- tolower(billboard2010Lyrics$track.name)
Top100of2010$track.name <- tolower(Top100of2010$track.name)

billboard2010Lyrics <- left_join(billboard2010Lyrics, Top100of2010, by = "track.name")
billboard2010Lyrics <- billboard2010Lyrics %>% inner_join(get_sentiments("bing"))

billboardLyrics$Billboard_Year <- "2022"
billboard2010Lyrics$Billboard_Year <- "2010"
totalData <- rbind(billboardLyrics, billboard2010Lyrics)

billboard2010LyricsSummary <- billboard2010Lyrics %>% summarise(positiveWords = sum(sentiment == "positive"), negativeWords = sum(sentiment == "negative"), posPercentage = positiveWords / (positiveWords + negativeWords), negPercentage = negativeWords / (positiveWords + negativeWords))
head(billboard2010LyricsSummary)
```


Combine Two Summaries From 2010 and 2022 Billboard Hot 100 Songs Lyrics Analysis
```{r}
Year <- c(rep("2010" , 2) , rep("2022" , 2) )
Sentiment <- rep(c("Positive" , "Negative") , 2)
Percentage <- c(0.5228659, 0.4771341, 0.3147662, 0.6852338)
finalSummary <- data.frame(Year,Sentiment,Percentage)
```



Visualising the share of negative words in lyrics in 2010 and 2022
```{r}
finalSummaryPlot <- finalSummary %>% 
  ggplot(aes(x = as.factor(Year), y = Percentage, fill = Sentiment)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  theme(panel.background = element_rect(fill = '#121212', color = '#121212'),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y = element_text(color="white", 
                                   size=12, family = "Arial Bold"), 
        axis.text.x = element_text(color="white", 
                                   size=12, family = "Arial Bold"), 
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = , l = 0))) +
  xlab("Year") + ylab("Percentage") + theme(text = element_text(color = "white", size = 15, family = "Arial Bold"),  plot.margin = margin(1, 1, 1, 1, "cm")) +
  theme(plot.background = element_rect(fill = "#121212"), legend.background = element_rect(fill = "#121212") ) +
  scale_fill_manual(values = c("Negative" = "#5727b0", "Positive" = "#ffea00"))

finalSummaryPlot
```



```{r}

allOut2020s <- get_playlist_audio_features("Vincent Balatero", "3aWW9eDfPZOy11QUMpqm0D")
valencePopularity <- allOut2020s %>% ggplot(aes(x = acousticness , y = track.popularity)) + geom_point() + geom_smooth()
valencePopularity
```


